---
title: "Team Project Multivariate Analysis"
author: "Adrian White, Cesar Conejo, Xavier Bryant"
date: "11/14/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load data, functions and libraries
load('rda/clinical_trial_complete.rda')
source("scripts/functions/graph_functions.R")
X <- clinical_trial_complete
# Remove the file
rm(clinical_trial_complete)

color_1 <- "deepskyblue2"
```

### Team members

* Adrian White
* Cesar Conejo: 100443596 (Representative member)
* Xavier Bryant: 100445659


### Introduction data set

We have selected the CRASH-2 data set for our project. It describes the outcome of a randomized controlled trial and economic valuation of the effects of tranexamic acid on death, vascular occlusive events and transfusion requirement in bleeding trauma patients. Tranexamic acid reduces bleeding in trauma patients undergoing surgery, but is an expensive treatnment option. The trial's objective was to assess the effects and cost effectiveness of an early administration of this medication. 

Participants of the study were adults with, or at risk of, significant bleeding within 8 hours of injury. Sample randomization was determined by the allocation of an eight digit sequence randomly generated by a computer. Patients and staff were masked to treatment allocation of the tranexamic acid. Health outcome was measured by number of life-years (LYs) gained, while cost data was obtained from the hospitals and measured by international collars ($) per LY. The trial was undertaken by two hundred and seventy-four hospitals world wide.


Some interesting links:

* https://pubmed.ncbi.nlm.nih.gov/23477634/
* https://www.thelancet.com/journals/lancet/article/PIIS0140-6736(10)60835-5/fulltext


info available in [Vanderbilt Biostatistics](http://biostat.mc.vanderbilt.edu/wiki/pub/Main/DataSets/crash2.html)


### Summary variables in the data set

Adrian/xavier. You can send me the columns with this info. By the moment, I am using this ones:


* entryid: (Numerical) Unique Numbers for Entry Forms

* sex: (Boolean) The sex of the patient

* age : (Numerical) Age of the patient

* injurytime: (Numerical) Hours since injury

* injurytype:	(Categorical) Type of injury {Blunt, Penetrating, Blunt and Penetrating}

* sbp: (Numerical) Systolic Blood Pressure

* rr: (Numerical) Respiratory Rate

* cc: (Numerical) Central Capillary Refille Time

* hr: (Numerical) Heart Rate

* icause: (Categorical) Cause of death

* istatus: (Categorical) Status of patient, if alive

* icondition: (Categorical) Condition of patient, if alive

* ndaysicu: (Numerical) Number of days in ICU 

* btransf: (Boolean) Blood Products Transfusion

* ncell: (Numerical) Number of Units of Red Call Products Transfused

* nplasma: (Numerical) Number of Units of Fresh Frozen Plasma Transfused

* nplatelets: (Numerical) Number of Units of Platelets Transfused

* ncryo: (Numerical) Number of Units of Cryoprecipitate Transfused

* bvii:	(Bloolean) Recombinant Factor VIIa Given


### Summary and Graphical display

A review of the structure of the dataset is the following:

```{r echo=FALSE}
str(X)
```

\newpage

A summary of the values in the data set are:

```{r echo=FALSE}
summary(X)
```

Finally, the list of different values by column is the following:

```{r echo=FALSE}
apply(X, 2, function(x) length(unique(x)))
```

Some visualizations of the variables is:

```{r echo=FALSE}
plothist(col_name = "age", df = X, color_1)
```

```{r echo=FALSE}
plothist(col_name = "injurytime", df = X, color_1)
```

```{r echo=FALSE}
plothist(col_name = "sbp", df = X, color_1)
```

```{r echo=FALSE}
plothist(col_name = "rr", df = X, color_1)
```

