---
title: "Team Project Multivariate Analysis"
author: "Adrian White, Cesar Conejo, Xavier Bryant"
date: "11/14/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load data, functions and libraries
load('rda/clinical_trial_complete.rda')
source("scripts/functions/graph_functions.R")
X <- clinical_trial_complete
# Remove the file
rm(clinical_trial_complete)

color_1 <- "deepskyblue2"
```

### Team members

* Adrian White
* Cesar Conejo: 100443596 (Representative member)
* Xavier Bryant: 100445659


### Introduction data set

Adrian resume about the CRASH2 trial in general:

Some interesting links:

* https://pubmed.ncbi.nlm.nih.gov/23477634/
* https://www.thelancet.com/journals/lancet/article/PIIS0140-6736(10)60835-5/fulltext


info available in [Vanderbilt Biostatistics](http://biostat.mc.vanderbilt.edu/wiki/pub/Main/DataSets/crash2.html)


### Summary variables in the data set

Adrian/xavier. You can send me the columns with this info. By the moment, I am using this ones:


* entryid: Unique Numbers for Entry Forms

* sex:

* ageP	

* injurytime: Hours Since Injury

* injurytype:	

* sbp: Systolic Blood Pressure

* rr: Respiratory Rate

* cc: Central Capillary Refille Time

* hr: Heart Rate

* btransf: Blood Products Transfusion

* ncell: Number of Units of Red Call Products Transfused

* nplasma: Number of Units of Fresh Frozen Plasma Transfused

* nplatelets: Number of Units of Platelets Transfused

* ncryo: Number of Units of Cryoprecipitate Transfused

* bvii:	Recombinant Factor VIIa Given

* boxid: Treatment Box Number

* packnum: Treatment Pack Number


### Summary and Graphical display

A review of the structure of the dataset is the following:

```{r echo=FALSE}
str(X)
```

\newpage

A summary of the values in the data set are:

```{r echo=FALSE}
summary(X)
```

Finally, the list of different values by column is the following:

```{r echo=FALSE}
apply(X, 2, function(x) length(unique(x)))
```

Some visualizations of the variables is:

```{r echo=FALSE}
plothist(col_name = "age", df = X, color_1)
```

```{r echo=FALSE}
plothist(col_name = "injurytime", df = X, color_1)
```

```{r echo=FALSE}
plothist(col_name = "sbp", df = X, color_1)
```

```{r echo=FALSE}
plothist(col_name = "rr", df = X, color_1)
```

